Experience Report: Haskell in Computational Biology

### 2012-03-09 mrfy_experience_report.tex
more math

These non-local interactions cause what had been a simple hidden Markov model 
to become a Markov random field.
SMURF~\cite{Menke:2010ti} and SMURFLite~\cite{Daniels:2012}, our previous work,
attempted to compute the 
optimal parse of a protein sequence onto the directed graph representing a 
Markov random field using multidimensional dynamic programming.
SMURF's 
computational complexity quickly exploded with complex protein structures, 
while SMURFLite limited its computational complexity by also limiting the 
complexity of protein structure it would understand.
The algorithm underlying 
MRFy grew out of an understanding of these limitations.

>>>>

These non-local interactions cause what had been a simple hidden Markov model 
to become a Markov random field.
SMURF~\cite{Menke:2010ti} and SMURFLite~\cite{Daniels:2012}, our previous work,
attempted to compute the 
optimal parse of a protein sequence onto the directed graph representing a 
Markov random field using multidimensional dynamic programming.
Mathematically, we can represent the probability of a query sequence 
$x_{1}..x_{n}$ given a Markov random field with an additional term in the
Viterbi relations.
Specifically, each term $V_{j}^{\prime M}(i)$ becomes
$W_{j}^{M}(i) = V_{j}^{\prime M}(i) - \log Pr(x_{i} | x_{\pi(j)})$, where
$x_{\pi(j)}$ is the amino acid in the query sequence that will be placed
in the beta-strand position paired with position $j$.
This nonlocal dependency prevents the problem from exhibiting optimal substructure,
preventing dynamic programming from solving the problem quickly.
SMURF's 
computational complexity quickly exploded with complex protein structures, 
while SMURFLite limited its computational complexity by also limiting the 
complexity of protein structure it would understand.
The algorithm underlying 
MRFy grew out of an understanding of these limitations.
